PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-24 paper/sec-introduction.tex
Introduction revised a little

The two respects are intertwined: PDFs provide a way of obtaining a prediction
for a given process in terms of other processes, therefore the accuracy and 
precision with which PDFs are able to make predictions crucially depend on 
the accuracy of the perturbative computation used in the PDF determination
(matched to the precision of the data). Electroweak (including combined QCD--EW) 
effects should therefore naturally be included in a PDF determination. The 
reason is twofold. First, one expects NNLO QCD and NLO EW corrections to be of 
comparable size because, at the EW scale, the QCD and EW running couplings 
become similar, $\alphas^2\sim \alpha$. If NNLO QCD corrections are included
by default in PDF determinations, NLO EW corrections should be
taken into account as well. Second, the virtual exchange of soft or 
collinear weak bosons leads to Sudakov 
logarithms~\cite{Denner:2000jv,Denner:2001gw},
which can make the coefficients of the EW series grow faster than 
their QCD counterparts. This behaviour is relevant in
phase-space regions associated with large mass scales (roughly of the order
of the \si{\tera\electronvolt}), where several LHC data sets (e.g.\ Z-boson
transverse-momentum distributions) enter both the determination of PDFs and the search
for new physics.

The systematic inclusion of EW corrections in a PDF determination entails
the solution of two separate problems. First, a problem of efficiency:
fast-interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are a fundamental ingredient
in any PDF fit, where the evaluation of the hadronic cross section needs to 
be performed a large number of times. While two default formats exist for
these grids, \textsc{APPLgrid}~\cite{Carli:2010rw} and
\textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}, none of them supports the inclusion of EW
corrections nor the interface to a Monte Carlo generator accurate to 
NLO QCD+EW\@. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such (e.g.\ the $t$-channel photon-induced component in
neutral-current Drell--Yan, which is not a separate processes beyond leading
order) or of just a part of the EW effects (e.g.\ multiple-photon
radiation from light particles in the final state of neutral- or charged-current
Drell--Yan, especially with electrons). Be that as it may, if EW effects are
systematically included in theoretical predictions, they should not be subtracted
from experimental results, otherwise they will be double counted.

@@

The paper is organised as follows. In section~\ref{sec:pineappl} we introduce
\textsc{PineAPPL}, we describe its features, we illustrate its operation, and we
assess its performance. In section~\ref{sec:results} we validate
\textsc{PineAPPL} and demonstrate its capabilities by computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW, for a representative
set of processes, usually included in PDF fits, for which EW corrections may
have a sizeable effect on the accuracy of the theoretical predictions.
In section~\ref{sec:doublecounting} we try to detail in a more
comprehensive manner the second problem sketched above, the solution of which,
however, remains beyond the scope of the current work. We provide our
conclusions and an outlook in section~\ref{sec:conclusion}. Examples of
installation and usage of \textsc{PineAPPL} are provided in
appendix~\ref{app:pineappl}; appendix~\ref{app:lumis} collects the parton
luminosities for each process considered in section~\ref{sec:results}; and
appendix~\ref{app:add_plots} complements some of the results presented
in section~\ref{sec:results}.

>>>>

The two respects are intertwined, as they both concur to determine the accuracy
of the theoretical predictions that are matched to the precision of the data.
Electroweak (including combined QCD--EW) effects should therefore naturally be
incorporated in calculations for LHC processes. The reason is twofold. First,
one expects NNLO QCD and NLO EW corrections to be of
comparable size because, at the EW scale, the QCD and EW running couplings 
become similar, $\alphas^2\sim \alpha$. If NNLO QCD corrections are included
by default in the computation, NLO EW corrections should be
taken into account as well. Second, the virtual exchange of soft or 
collinear weak bosons leads to Sudakov 
logarithms~\cite{Denner:2000jv,Denner:2001gw},
which can make the coefficients of the EW series grow faster than 
their QCD counterparts. This behaviour is relevant in
phase-space regions associated with large mass scales (roughly of the order
of the \si{\tera\electronvolt}), where several LHC data sets (e.g.\ Z-boson
transverse-momentum distributions) enter both the determination of PDFs and the search
for new physics.

The systematic inclusion of EW corrections in precision computations for
LHC processes entails
the solution of two separate problems. First, a problem of efficiency:
fast-interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are essential
whenever the evaluation of the hadronic cross section needs to 
be performed a large number of times, as is the case in the evaluation of
scale variations or of PDF fits. While two default formats exist for
these grids, \textsc{APPLgrid}~\cite{Carli:2010rw} and
\textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}, none of them supports the inclusion of EW
corrections nor the interface to a Monte Carlo generator accurate to 
NLO QCD+EW\@. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such (e.g.\ the $t$-channel photon-induced component in
neutral-current Drell--Yan, which is not a separate processes beyond leading
order) or of just a part of the EW effects (e.g.\ multiple-photon
radiation from light particles in the final state of neutral- or charged-current
Drell--Yan, especially with electrons). Be that as it may, if EW effects are
systematically included in theoretical predictions, they should not be subtracted
from experimental results, otherwise they will be double counted.

@@

The paper is organised as follows. In section~\ref{sec:pineappl} we introduce
\textsc{PineAPPL}, we describe its features, we illustrate its operation, and we
assess its performance. In section~\ref{sec:results} we validate
\textsc{PineAPPL} and demonstrate its capabilities by computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW, for a representative
set of LHC processes for which EW corrections may
have a sizeable effect on the accuracy of the theoretical predictions.
In section~\ref{sec:doublecounting} we try to detail in a more
comprehensive manner the second problem sketched above, the solution of which,
however, remains beyond the scope of the current work. We provide our
conclusions and an outlook in section~\ref{sec:conclusion}. Examples of
installation and usage of \textsc{PineAPPL} are provided in
appendix~\ref{app:pineappl}; appendix~\ref{app:lumis} collects the parton
luminosities for each process considered in section~\ref{sec:results}; and
appendix~\ref{app:add_plots} complements some of the results presented
in section~\ref{sec:results}.
